---
  title: "SOP correction for the assessment of Megrim (*Lepidorhombus whiffiagonis*) in Divisions 7.b-k, 8.a, 8.b, and 8.d"
author: 
  - Miren Altuna-Etxabe 
- Santiago Cerviño
- Dorleta García
date: today
bibliography: docs/ref.bib 
format: 
  html:
  toc: true
toc-depth: 5
css: docs/styles.css
fig-cap: true
fig-cap-location: bottom
self-contained: true
#   code-fold: true
# pdf:
# toc: true
# toc-depth: 5
# css: styles.css
# number-sections: true
# fig-pos: 'H'
# fig-cap: true
# fig-cap-location: bottom
# code-block-font-size: 10pt
# code-block-wrap: true
# documentclass: article
# keep-tex: true
# pdf-engine: xelatex
# mainfont: "Arial"
# block-headings: true
# include-in-header: preamble.tex
# include-before-toc: | 
# \clearpage
# include-after-toc: | 
# \clearpage
execute:
  crossref: true
---
  \newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newpage

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyr)
library(ggplot2)
library(readxl)
library(dplyr)
library(FLCore)

last.yr = 2025
```
There are SOP problems in the catch, landings and discards of the last year data (@fig-sop).

```{r}
#| echo: false
#| label: fig-sop
#| fig-cap: "Percentage differences between the total reported catch, landings and discards and the sum of the product of numbers-at-age and weigth-at-age."
#| message: false
#| warning: false

# load original data without removing stock@catch.n['1',as.character(1984:2000)]
load("Input/bootstrap/data/stock/meg78_stock.RData")

# CATCH SOP problem
nwt <- c(apply(stock@catch.n * stock@catch.wt, 2, function(x) sum(x, na.rm = TRUE)))
total <- c(stock@catch)
diff_c <- total / nwt
diff_c_prc <- as.numeric((1 - diff_c) * 100)

catch.sop <- data.frame(
  year = dimnames(stock@catch)$year,
  nwt = nwt,
  Total = total,
  Diff = diff_c,
  Diff_prc = round(diff_c_prc, 2)
)

# LANDINGS SOP problem
nwt <- c(apply(stock@landings.n * stock@landings.wt, 2, function(x) sum(x, na.rm = TRUE)))
total <- c(stock@landings)
diff_l <- total / nwt
diff_l_prc <- as.numeric((1 - diff_l) * 100)

landings.sop <- data.frame(
  year = dimnames(stock@landings)$year,
  nwt = nwt,
  Total = total,
  Diff = diff_l,
  Diff_prc = round(diff_l_prc, 2)
)

# DISCARDS SOP problem
nwt <- c(apply(stock@discards.n * stock@discards.wt, 2, function(x) sum(x, na.rm = TRUE)))
total <- c(stock@discards)
diff_d <- total / nwt
diff_d_prc <- as.numeric((1 - diff_d) * 100)

discards.sop <- data.frame(
  year = dimnames(stock@discards)$year,
  nwt = nwt,
  Total = total,
  Diff = diff_d,
  Diff_prc = round(diff_d_prc, 2)
)

# SOP summary table
sop <- data.frame(
  year = dimnames(stock@catch)$year,
  catch = round(diff_c_prc, 2),
  landings = round(diff_l_prc, 2),
  discards = round(diff_d_prc, 2)
)

sop[sop == -Inf] <- ""

sop_long <- sop %>%
  mutate(
    Catch = as.numeric(catch),
    Landings = as.numeric(landings),
    Discards = as.numeric(discards)) %>%
 pivot_longer(cols = c(Catch, Landings, Discards),
               names_to = "type",
               values_to = "value") %>% select(year, type, value)

sop_long$type    <- factor(sop_long$type, levels = c("Catch", "Landings", "Discards"))

sop_long <- sop_long %>% filter(year == last.yr)

# SOP summary PLOT
ggplot(sop_long, aes(x = factor(year), y = value, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year", y = "Difference (%)", fill = "Type") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust =1))
```

We analyzed the input data submitted by each country in the last year. Our evaluation shows that the product of numbers-at-age and weight-at-age for both landings and discards does not consistently match the reported totals for these categories at the country level. These discrepancies are visualized in @fig-input_ctry.

In this case, we also analize the discrepancies in 2022 because there was a error in Belgium landings data and we recalculate the SOP error and correct them. The number of individuals used in the assessment as landings for Belgium in 2022 was incorrect, as it actually corresponded to the sum of landings and discards.

This inconsistency indicates that the differences observed in the total SOP originate from the input data submitted by individual countries.

```{r}
#| echo: false
#| label: fig-input_ctry
#| fig-cap: "Differences in percentage between the product of numbers-at-age and weight-at-age for both landings and discards and the reported totals for these categories at the country level over the last three years. Countries: BE = Belgium, FR = France, IR = Ireland, SP = Spain, UK = United Kingdom."
#| message: false
#| warning: false

LD_cntry <- read_excel("Input/Ctry_info/TLD_cntry.xlsx")
LD_age_cntry <- read_excel("Input/Ctry_info/LD_age_cntry.xlsx") # Original nind and wt from intercatch

LD_cntry_wt_ind <- LD_age_cntry %>%
  pivot_longer(cols = `1`:`10`, names_to = "age", values_to = "value")%>%
  pivot_wider(names_from = param, values_from = value) %>%
  select(year, country, catch.cat, age, wt, nind)

# calculate the difference among total landings/discards and the sum of the nind*wt by country
LD_age_cntry_long <- LD_cntry_wt_ind %>%
  mutate(twt = wt*nind) %>%
  group_by(year, country, catch.cat) %>% 
  summarise(total_wtnind = sum(twt), .groups = "drop") %>%
  select(year, country, catch.cat, total_wtnind)

LD_cntry_info <- full_join(LD_age_cntry_long, LD_cntry, by = c("year", "country", "catch.cat"))

LD_cntry_info <- LD_cntry_info %>%
  mutate(diff = round(as.numeric(LD_cntry_info$total/LD_cntry_info$total_wtnind),2),
         diff_prc = (1 - as.numeric(diff)) * 100,
         year = as.factor(year),
         catch.cat = case_when(
                      catch.cat == "discards" ~ "Discards",
                      catch.cat == "landings" ~ "Landings",
                      TRUE ~ catch.cat  # mantiene los demás valores sin cambios
                    )) %>% filter(year %in% c(2022, last.yr))

LD_cntry_info$catch.cat <- factor(LD_cntry_info$catch.cat, levels = c("Landings", "Discards"))

ggplot(LD_cntry_info, aes(x = country, y = diff_prc, fill = factor(year))) +
                    geom_bar(stat = "identity", position = "dodge") +
                    facet_grid(. ~ catch.cat) +
                    labs(
                      x = "Country",
                      y = "Difference (%)",
                      fill = "Year") +
                    theme_bw() +
                    theme(axis.text.x = element_text(hjust = 0.5),
                          strip.text = element_text(size = 16))
```

To address the SOP-related issues affecting the catch, landings, and discards data, the SOP corrections in numbers-at-age level were applied at landings and discards level separately at country level using the mean weight-at-age across all countries to ensure consistency with total reported values. SOP issues in catch data were addressed by summing the updated numbers of individuals from landings and discards, and calculating the total catch weight using a weighted average of the weight-at-age data from landings and discards.

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

# At country level + total wt
landings.wt <- as.data.frame(stock@landings.wt)
landings.wt<- landings.wt %>% 
  mutate(wt = 1000*data, catch.cat = "landings") %>%
  select(year, catch.cat, age, wt)

discards.wt <- as.data.frame(stock@discards.wt)
discards.wt<- discards.wt %>%
  mutate(wt = 1000*data, catch.cat = "discards") %>%
  select(year, catch.cat, age, wt)

LD_total_wt <- rbind(landings.wt, discards.wt)
LD_cntry_wt_ind <- LD_cntry_wt_ind %>% mutate(age=as.numeric(age))

LD_joined <- LD_cntry_wt_ind %>% mutate(nind = if_else(catch.cat == "discards" & age == 9 |catch.cat == "discards" & age == 10, 0, nind)) %>%
  inner_join(
    LD_total_wt %>% select(year, catch.cat, age, wt),
    by = c("year", "catch.cat", "age"),
    suffix = c("_cntry", "_total")
  )

LD_joined_wt_total <- LD_joined %>%
  select(year, country, catch.cat, age, wt_total, nind)  %>%
  mutate(totalwt = wt_total*nind) %>%
  group_by(year, country, catch.cat) %>% 
  summarise(total_totalwt = sum(totalwt), .groups = "drop") %>%
  select(year, country, catch.cat, total_totalwt)

LD_joined_wt_total <- full_join(LD_joined_wt_total, LD_cntry, by = c("year", "country", "catch.cat"))
LD_joined_wt_total$diff <-round(LD_joined_wt_total$total/LD_joined_wt_total$total_totalwt, 2)
LD_joined_wt_total$multiplier <- LD_joined_wt_total$total/LD_joined_wt_total$total_totalwt

ctry_sop_dif <- LD_joined_wt_total

cols_to_multiply <- as.character(1:10)
LD_ctry <- LD_age_cntry  %>% filter(param == "nind") %>%
  left_join(LD_joined_wt_total %>% select(year, country, catch.cat, multiplier),
            by = c("year", "country", "catch.cat")) %>%
  mutate(across(all_of(cols_to_multiply), ~ .x * multiplier))

LD_ctry <- LD_ctry %>% select(-multiplier)
LD_ctry <- LD_ctry %>%
  group_by(year, catch.cat) %>%
  summarise(across(`1`:`10`, sum), .groups = "drop") %>% 
  mutate( `9` = if_else(catch.cat == "discards", 0, `9`),
          `10` = if_else(catch.cat == "discards", 0, `10`))

for (year in last.yr) {
  stock@landings.n[,ac(year)] <- as.numeric(LD_ctry[LD_ctry$year==year & LD_ctry$catch.cat=="landings", -c(1:2)])
  stock@discards.n[,ac(year)] <- as.numeric(LD_ctry[LD_ctry$year==year & LD_ctry$catch.cat=="discards", -c(1:2)])
}


# Catch nind
for (year in last.yr) {
  stock@catch.n[,ac(year)] <- stock@landings.n[,ac(year)] + stock@discards.n[,ac(year)]
}

# wt
landings.n <- as.data.frame(stock@landings.n) %>% filter(year == last.yr) %>% select(age, year, data) %>% rename(n_landings = data)
landings.wt <- landings.wt %>% filter(year == last.yr) %>% select(age, year, wt) %>% rename(wt_landings = wt)

discards.n <- as.data.frame(stock@discards.n) %>% filter(year == last.yr) %>% select(age, year, data)%>% rename(n_discards = data)
discards.wt <- discards.wt %>% filter(year == last.yr) %>% select(age, year, wt) %>% rename(wt_discards = wt)

df <- landings.n %>%
  left_join(landings.wt, by = c("year", "age")) %>%
  left_join(discards.n, by = c("year", "age")) %>%
  left_join(discards.wt, by = c("year", "age"))

df <- df %>% 
  mutate(
    n_landings = replace_na(n_landings, 0),
    n_discards = replace_na(n_discards, 0),
    wt_discards = replace_na(wt_discards, 0),
    WT = case_when(
      (n_landings + n_discards) > 0 ~
        (n_landings * wt_landings + n_discards * wt_discards) /
        (n_landings + n_discards),
      TRUE ~ NA_real_
    )
  )

df_wide <- df %>%
  select(year, age, WT) %>%
  pivot_wider(names_from = age, values_from = WT)

for (year in last.yr) {
  stock@catch.wt[,ac(year)] <- as.numeric(df_wide[df_wide$year==year, -1])/1000
}

save(stock, file="Input/bootstrap/data/stock/meg78_stock_sop_corrected.RData")
```

```{r}

## To correct 2022 BE error

landings.wt <- as.data.frame(stock@landings.wt)
landings.wt<- landings.wt %>% 
  mutate(wt = 1000*data, catch.cat = "landings") %>%
  select(year, catch.cat, age, wt)

discards.wt <- as.data.frame(stock@discards.wt)
discards.wt<- discards.wt %>%
  mutate(wt = 1000*data, catch.cat = "discards") %>%
  select(year, catch.cat, age, wt)

LD_total_wt <- rbind(landings.wt, discards.wt)
LD_cntry_wt_ind <- LD_cntry_wt_ind %>% mutate(age=as.numeric(age))

LD_joined <- LD_cntry_wt_ind %>% mutate(nind = if_else(catch.cat == "discards" & age == 9 |catch.cat == "discards" & age == 10, 0, nind)) %>%
  inner_join(
    LD_total_wt %>% select(year, catch.cat, age, wt),
    by = c("year", "catch.cat", "age"),
    suffix = c("_cntry", "_total")
  )

LD_joined_wt_total <- LD_joined %>%
  select(year, country, catch.cat, age, wt_total, nind)  %>%
  mutate(totalwt = wt_total*nind) %>%
  group_by(year, country, catch.cat) %>% 
  summarise(total_totalwt = sum(totalwt), .groups = "drop") %>%
  select(year, country, catch.cat, total_totalwt)

LD_joined_wt_total <- full_join(LD_joined_wt_total, LD_cntry, by = c("year", "country", "catch.cat"))
LD_joined_wt_total$diff <-round(LD_joined_wt_total$total/LD_joined_wt_total$total_totalwt, 2)
LD_joined_wt_total$multiplier <- LD_joined_wt_total$total/LD_joined_wt_total$total_totalwt

ctry_sop_dif <- LD_joined_wt_total

cols_to_multiply <- as.character(1:10)
LD_ctry <- LD_age_cntry  %>% filter(param == "nind") %>%
  left_join(LD_joined_wt_total %>% select(year, country, catch.cat, multiplier),
            by = c("year", "country", "catch.cat")) %>%
  mutate(across(all_of(cols_to_multiply), ~ .x * multiplier))

LD_ctry <- LD_ctry %>% select(-multiplier)
LD_ctry <- LD_ctry %>%
  group_by(year, catch.cat) %>%
  summarise(across(`1`:`10`, sum), .groups = "drop") %>% 
  mutate( `9` = if_else(catch.cat == "discards", 0, `9`),
          `10` = if_else(catch.cat == "discards", 0, `10`))


for (year in c(2022)) {
  stock@landings.n[,ac(year)] <- as.numeric(LD_ctry[LD_ctry$year==year & LD_ctry$catch.cat=="landings", -c(1:2)])
  stock@discards.n[,ac(year)] <- as.numeric(LD_ctry[LD_ctry$year==year & LD_ctry$catch.cat=="discards", -c(1:2)])
}


# Catch nind
for (year in c(2022)) {
  stock@catch.n[,ac(year)] <- stock@landings.n[,ac(year)] + stock@discards.n[,ac(year)]
}

# wt
landings.n <- as.data.frame(stock@landings.n) %>% filter(year == 2022) %>% select(age, year, data) %>% rename(n_landings = data)
landings.wt <- landings.wt %>% filter(year == 2022) %>% select(age, year, wt) %>% rename(wt_landings = wt)

discards.n <- as.data.frame(stock@discards.n) %>% filter(year == 2022) %>% select(age, year, data)%>% rename(n_discards = data)
discards.wt <- discards.wt %>% filter(year == 2022) %>% select(age, year, wt) %>% rename(wt_discards = wt)

df <- landings.n %>%
  left_join(landings.wt, by = c("year", "age")) %>%
  left_join(discards.n, by = c("year", "age")) %>%
  left_join(discards.wt, by = c("year", "age"))

df <- df %>%
  mutate(
    n_landings = replace_na(n_landings, 0),
    n_discards = replace_na(n_discards, 0),
    wt_discards = replace_na(wt_discards, 0),
    WT = case_when(
      (n_landings + n_discards) > 0 ~
        (n_landings * wt_landings + n_discards * wt_discards) /
        (n_landings + n_discards),
      TRUE ~ NA_real_
    )
  )

df_wide <- df %>%
  select(year, age, WT) %>%
  pivot_wider(names_from = age, values_from = WT)

  stock@catch.wt[,ac(2022)] <- as.numeric(df_wide[df_wide$year==2022, -1])/1000
  
  save(stock, file="Input/bootstrap/data/stock/meg78_stock_sop_BE_corrected.RData")

```

Chek if the sop problem in the last year is solved or not

```{r}
#| echo: false
#| label: fig-sop
#| fig-cap: "Percentage differences between the total reported catch, landings and discards and the sum of the product of numbers-at-age and weigth-at-age."
#| message: false
#| warning: false

# CATCH SOP problem
nwt <- c(apply(stock@catch.n * stock@catch.wt, 2, function(x) sum(x, na.rm = TRUE)))
total <- c(stock@catch)
diff_c <- total / nwt
diff_c_prc <- as.numeric((1 - diff_c) * 100)

catch.sop <- data.frame(
  year = dimnames(stock@catch)$year,
  nwt = nwt,
  Total = total,
  Diff = diff_c,
  Diff_prc = round(diff_c_prc, 2)
)

# LANDINGS SOP problem
nwt <- c(apply(stock@landings.n * stock@landings.wt, 2, function(x) sum(x, na.rm = TRUE)))
total <- c(stock@landings)
diff_l <- total / nwt
diff_l_prc <- as.numeric((1 - diff_l) * 100)

landings.sop <- data.frame(
  year = dimnames(stock@landings)$year,
  nwt = nwt,
  Total = total,
  Diff = diff_l,
  Diff_prc = round(diff_l_prc, 2)
)

# DISCARDS SOP problem
nwt <- c(apply(stock@discards.n * stock@discards.wt, 2, function(x) sum(x, na.rm = TRUE)))
total <- c(stock@discards)
diff_d <- total / nwt
diff_d_prc <- as.numeric((1 - diff_d) * 100)

discards.sop <- data.frame(
  year = dimnames(stock@discards)$year,
  nwt = nwt,
  Total = total,
  Diff = diff_d,
  Diff_prc = round(diff_d_prc, 2)
)

# SOP summary table
sop <- data.frame(
  year = dimnames(stock@catch)$year,
  catch = round(diff_c_prc, 2),
  landings = round(diff_l_prc, 2),
  discards = round(diff_d_prc, 2)
)

sop[sop == -Inf] <- ""

sop_long <- sop %>%
  mutate(
    Catch = as.numeric(catch),
    Landings = as.numeric(landings),
    Discards = as.numeric(discards)) %>%
 pivot_longer(cols = c(Catch, Landings, Discards),
               names_to = "type",
               values_to = "value") %>% select(year, type, value)

sop_long$type    <- factor(sop_long$type, levels = c("Catch", "Landings", "Discards"))

sop_long <- sop_long %>% filter(year == c(last.yr, 2022))

# SOP summary PLOT
ggplot(sop_long, aes(x = factor(year), y = value, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year", y = "Difference (%)", fill = "Type") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust =1))
```